Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-08-12 paper/ms.tex
new full version

This problem can be tackeld with Monte Carlo simulations for the
radiative transfer, however the complexity of the relevant physical
processes makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift.

@@

In the next subsections we describe in detail the four key
elements of this workflow. First, we present the observations we take
as a benchmark. Second, we describe the main characteristics of the
N-body simulation and the halo catalogs we use. Third, we recount the
important parameters of the simplified model that we use to populate
the halo catalogs with LAEs. Finally we describe some of the
statistical tests we adopt to compare observations and mocks.

@@

The first observational constraint we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey
conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera,
which has a field of view covering $34\times 27$ arcmin, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$.  The narrow
band filter used in the survey is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors reported a
total  $2161$  LAEs with an observed equivalent width larger than $190$\AA 
over a total survey area of $2.42$ deg$^{2}$ that includes 12 subfields, 
this corresponds to average surface number density of $0.20\pm 0.01$
arcmin$^{-2}$.

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs (7 subfields), this
field has been known to harbor a region with a large density excess of
galaxies. The second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a
total of $0.58$ deg$^2$ and $386$ LAEs (3 subfields). The third and
fourth fields are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and
$196$ LAEs (1 subfield), and the field around the Great Observatory
Optical Deep Survey  (GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs (1
subfield).

@@

We assume that a dark matter halo can only host one
detectable LAE at most.  There are three parameters that
decide whether a halo host a LAE: the lower and upper bounds for the
mass range $M_{\rm min}< M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm occ}$ of such halos that host a detectable LAE. The reader
must keep in mind that the physical interpreation of the occupation
fraction $f_{\rm occ}$ convolves two phenomena: the presence of a star
forming galaxy in a halo and its detectability as a LAE.  We do not
peform a explicit modeling for LAE detectability. Our model does not
assign a luminosity or escape fraction to each LAE.  We are only
interested in constraining the halo mass range hosting detectable LAEs
under the conditions defined by \cite{Yamada2012}.

@@

The second constraint  comes from the angular correlation function (ACF)
for all the models having the $15$ mock surveys consistent with
observations . The ACF is computed  ( using 
the Davis \& Peebles  estimator \citep{Davis1983} ) only on the densest
subfield in all the 15 mock surveys corresponding to the SSA22
region. We then take the mean ACF over the 15  mock surveys  for
each model as well as its the standard deviation to include
the effect of cosmic variance.

We finally compare the observational ACF and the mean mock  ACF of the models   
in therms of the angular correlation lenght ($\theta_{0}$) by fitting 
them to a power-law function:

@@

Both mock and observational ACF are derived by a least
square minimization procedure. The observational errors of 
the ACF as well as the standard deviation of the mean modelled ACFs  are
considered to compute the error in the fitted parameters to ensure ourselves
to include poissionian and cosmic-variance errors. 
We consider that a model is consistent with observations if the two
parameters $\beta$ and $\theta_0$ are equal within a $1$-$\sigma$ range.

@@

The second region corresponds to an intermediate range for the minimum
mass $10.5<\log_{10}M_{\rm min}<11.5$ where, regardless of the value of
the maximum mass $M_{\rm max}$, it is possible to tune the occupation
fraction $f_{\rm occ}$ to bring some of the mock observations into
good agreement with observations. In this region of parameter space
one can tfind two extreme kinds of models.  One kind where the
mass interval $\Delta M\equiv \log_{10} M_{\rm max} - \log_{10}M_{\rm
  min}$ is narrow with $\Delta M<1.0$ dex, others where the mass interval is 
extended $\Delta M>1.0$dex going up to the maximum halo mass
present in the simulation at that redshift, up to $\Delta M = 2.5$ dex
in some cases.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig6_mass.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig6_f_occ.pdf}
\end{center}
\caption{Planes $M_{\rm min}$ vs $M_{\rm max}$ (left), $M_{\rm min}$
  vs $f_{\rm occ}$ (rigth). The squares represent the models when both
  th constraints on the maximal number of  consistent mocks and the
  angular correlation function are included. 
  \label{fig:restriction_mock_and_f_occ_corr}} 
\end{figure*}

@@

One sceneario can be presented in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems. Due to the resonant nature of the
\ly line the probability of absortion  of \ly photons increases in
massive systems, producing high absorption of \ly photons but not of
continuum or other non-resonant lines
\citep{Laursen2009,ForeroRomero2011}.

@@

The preference for narrow $\Delta M$ ranges to host LAEs 
and the minoritary presence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun follows theoretical and observational
insights where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012,Shapley2003}.

@@

We also find that an ACF analysis on a single dataset is not enough to rule
out models with a high occupation fraction $f_{\rm occ}>0.3$, which
represent almost one fourth of our best models, coinciding with a wide
range in halo masses $\Delta M>1.0$ dex. These models cannot be
considered unfavorable based on a clustering analysis only.

\cite{Ouchi2010} presents an analysis of LAE observations in the
redshift interval $3.1<z<7.0$. At $z=3.1$ They get to an average
mass for the host dark matter halos of $M_{h}=2.9^{+24.0}_{-2.9}\times
10^{10}$ \hMsun with a corresponding duty cycle of $0.008\pm
0.03$. This broadly matches the expectations from the first family of
models, also summarized in Table \ref{table:firstfamily}. In
particular it seems to favor only the very low occupations fractions.

@@

Abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} work based on observational
results for Lyman Break Galaxies (LBG). In those studies the minimum halo
mass considered to be relevant in their analysis is
$10^{11.4}$\hMsun. For that mass, \cite{Behroozi2013a} report a
stellar mass mass around $(1.0\pm0.3)\times 10^{9.0}$
\hMsun, while their star formation rate is in the range $0.6\pm 0.2$
Msun yr$^{-1}$, which nevertheless are close to the lower bound of
values inferred for LAEs at high redshift
\citep{Gawiser2007,Nilsson2009,Pentericci2009}.

In our results, all the preferred models have a halo mass range lower
than the minimum of $M_{\rm min}<10^{11.4}$\hMsun considerend in
abundance matching at $z=3$. This shows that our results also confirm
the expectations of LAEs to be hosted by less massive halos than LBGs.
A detaile analysis of the spectral and photometric properties of LAEs
coupled to the kind of analysis performed in this paper can be a guide
in the study of the properties of low mass dark matter halos at
$z=3.1$, extending the capabilities of abundance matching methods.

\subsection{Possible caveats to our method}

... Single LAE in a halo.

... Mass definition for different halo finders.

@@

In this \documentname we look for constraints on the preferred mass
and occupation fraction of dark matter halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We perform this study
paying attention to the impact of cosmic variance in these results. We
build a large number of mock catalogs matching observational
geometries. The mocks are constructed from a N-body simulation
following a simple recipe to assign a single LAE to each halo. Only
a fraction $f_{\rm occ}$ of halos with a mass range  $M_{\rm
  min}<M_{\rm h}<M_{\rm   max}$ can host a LAE. We proceed with a
thourough exploration of the space of free parameters $M_{\rm min},
M_{\rm max}$ and $f_{\rm occ}$ to find mocks that are consistent with
two observational constraints: the surface number density and the
angular correlation function. Out of the initial $9000$ initial
combinations of parameters in the model we find $49$ arrangements
consistent with observations.

The most relevant result from our study is the following. The 
successful parameter arrangements present a wide range of possibilities for
DM hosting LAE hosts at $z=3.1$. This means that the available LAEs' spatial
information is not sufficient to put a tight constraint on the mass
and occupation fraction of DM halos hosting LAEs. In particular it is
not possible to put a strong constraint on the occupation
fraction. Such a wide range in solutions is facilitated by the large
dispersion in the statistics derived from the mocks. All the halo mass
ranges and occupation fractions deduced in previous analysis in
previous analysis \citep[i.e.][]{Gawiser2007} are a subset of the
models we find in this \documentname.

Nevertheless, the sucessful models can be split into three families
depending the mass range $\Delta M=\log_{10}M_{\rm max} -
\log_{10}M_{\rm min}$ and the occupation fraction $f_{\rm occ}$. The
first family is narrow both in $\Delta M$ and $f_{\rm occ}$, a second
family is only narrow in $\Delta M$ and the third is narrow in $f_{\rm
  occ}$ and broad in $\Delta M$.

A second results from our investigation is the existence
of a large number of models with a very narrow mass range $\Delta M<
0.1$ dex that are consistent in every aspect with the spatial
distribution of observed LAEs. In order to have these characteristics
it is required that halos around the mass $M_{\rm   max}$ for each
model become inefficient in hosting detectable LAEs. Two possible
ways to achieve this is having LAEs with a decreasing \ly escape
fraction with increasing mass or increasing screening effects by
neutral Hydrogen around the most massive systems
\citep{Laursen2009,ForeroRomero2011}.

A third aspect that stands out is the existence of a dominant family
of models with narrow $\Delta M<0.1$, as described before, and also a
low occupation fraction $f_{\rm occ}<0.3$. This gives support to the
common wisdom stating that only a small fraction of star forming
galaxies can be detected as strong LAEs.

We foresee that the new observations with new instruments (such as MUSE,
Hyper SuprimeCam and HETDEX) covering larger fields and a wider range
of luminosities will be key in imposing tighter constraints on the
properties of dark matter halos hosting LAEs.

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed. The authors also thank Peter Laursen, Paulina Lira and
Alvaro Orsi for helpful comments on the physical interpretation and
the presentation of our results.

>>>>

This problem can be tackled with Monte Carlo simulations for the
radiative transfer, however the complexity of the relevant physical
processes makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift.

@@

In the next subsections we describe in detail the four key
elements of this work-flow. First, we present the observations we take
as a benchmark. Second, we describe the main characteristics of the
N-body simulation and the halo catalogs we use. Third, we recount the
important parameters of the simplified model that we use to populate
the halo catalogs with LAEs. Finally we describe some of the
statistical tests we adopt to compare observations and mocks.

@@

The first observational constraint we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey
conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera,
which has a field of view covering $34\times 27$ arc-min, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$.  The narrow
band filter used in the survey is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors reported a
total  $2161$  LAEs with an observed equivalent width larger than $190$\AA 
over a total survey area of $2.42$ deg$^{2}$ that includes 12 sub-fields, 
this corresponds to average surface number density of $0.20\pm 0.01$
arcmin$^{-2}$.

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs (7 sub-fields), this
field has been known to harbor a region with a large density excess of
galaxies. The second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a
total of $0.58$ deg$^2$ and $386$ LAEs (3 sub-fields). The third and
fourth fields are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and
$196$ LAEs (1 sub-field), and the field around the Great Observatory
Optical Deep Survey  (GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs (1
sub-field).

@@

We assume that a dark matter halo can only host one
detectable LAE at most.  There are three parameters that
decide whether a halo host a LAE: the lower and upper bounds for the
mass range $M_{\rm min}< M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm occ}$ of such halos that host a detectable LAE. The reader
must keep in mind that the physical interpretation of the occupation
fraction $f_{\rm occ}$ convolves two phenomena: the presence of a star
forming galaxy in a halo and its detectability as a LAE.  We do not
peform a explicit modeling for LAE detectability. Our model does not
assign a luminosity or escape fraction to each LAE.  We are only
interested in constraining the halo mass range hosting detectable LAEs
under the conditions defined by \cite{Yamada2012}.

@@

The second constraint  comes from the angular correlation function (ACF)
for all the models having the $15$ mock surveys consistent with
observations . The ACF is computed  ( using 
the Davis \& Peebles  estimator \citep{Davis1983} ) only on the densest
sub-field in all the 15 mock surveys corresponding to the SSA22
region. We then take the mean ACF over the 15  mock surveys  for
each model as well as its the standard deviation to include
the effect of cosmic variance.

We finally compare the observational ACF and the mean mock  ACF of the models   
in therms of the angular correlation length ($\theta_{0}$) by fitting 
them to a power-law function:

@@

Both mock and observational ACF are derived by a least
square minimization procedure. The observational errors of 
the ACF as well as the standard deviation of the mean modeled ACFs  are
considered to compute the error in the fitted parameters to ensure ourselves
to include poissonian and cosmic-variance errors. 
We consider that a model is consistent with observations if the two
parameters $\beta$ and $\theta_0$ are equal within a $1$-$\sigma$ range.

@@

The second region corresponds to an intermediate range for the minimum
mass $10.5<\log_{10}M_{\rm min}<11.5$ where, regardless of the value of
the maximum mass $M_{\rm max}$, it is possible to tune the occupation
fraction $f_{\rm occ}$ to bring some of the mock observations into
good agreement with observations. In this region of parameter space
one can find two extreme kinds of models.  One kind where the
mass interval $\Delta M\equiv \log_{10} M_{\rm max} - \log_{10}M_{\rm
  min}$ is narrow with $\Delta M<1.0$ dex, others where the mass interval is 
extended $\Delta M>1.0$dex going up to the maximum halo mass
present in the simulation at that redshift, up to $\Delta M = 2.5$ dex
in some cases.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig6_mass.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig6_f_occ.pdf}
\end{center}
\caption{Planes $M_{\rm min}$ vs $M_{\rm max}$ (left), $M_{\rm min}$
  vs $f_{\rm occ}$ (right). The squares represent the models when both
   constraints on the maximal number of  consistent mocks and the
  angular correlation function are included. 
  \label{fig:restriction_mock_and_f_occ_corr}} 
\end{figure*}

@@

One scenario can be presented in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems. Due to the resonant nature of the
\ly line the probability of absorption  of \ly photons increases in
massive systems, producing high absorption of \ly photons but not of
continuum or other non-resonant lines
\citep{Laursen2009,ForeroRomero2011}.

@@

The preference for narrow $\Delta M$ ranges to host LAEs 
and the small presence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun follows theoretical and observational
insights where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012,Shapley2003}.

@@

We also find that an ACF analysis on a single data-set is not enough to rule
out models with a high occupation fraction $f_{\rm occ}>0.3$, which
represent almost one fourth of our best models, coinciding with a wide
range in halo masses $\Delta M>1.0$ dex. These models cannot be
considered unfavorable based on a clustering analysis only.

\cite{Ouchi2010} presents an analysis of LAE observations in the
redshift interval $3.1<z<7.0$. At $z=3.1$ They quote an average
mass for the host dark matter halos of $M_{h}=2.9^{+24.0}_{-2.9}\times
10^{10}$ \hMsun with a corresponding duty cycle of $0.008\pm
0.03$.  This broadly matches the expectations from the first family of
models, also summarized in Table \ref{table:firstfamily}. In
particular it seems to favor only the very low occupations
fractions.

We conclude this section by noting that the uncertainties on
the preferred masses in \citep{Gawiser07} and \cite{Ouchi2010} are
found with in the intervals $M_{\rm min}$ and $M_{\rm max}$ of our
models. Under the light of our model, this implies that detected LAEs
must be absent from the most massive dark matter halos at that
redshift.

@@

The abundance matching methods are based on observational results for Lyman Break Galaxies
\citep{Behroozi2013a,Behroozi2013b}  (LBG).  In the case of
\cite{Behroozi2013a} the minimum halo mass considered to be relevant
in their analysis is $10^{11.4}$\hMsun. They report  stellar mass mass
around $(1.0\pm0.3)\times 10^{9.0}$ \hMsun, while their star formation
rate is in the range $0.6\pm 0.2$ Msun yr$^{-1}$, which nevertheless
are close to the lower bound of values inferred for LAEs at high
redshift \citep{Gawiser2007,Nilsson2009,Pentericci2009}.

In our results, all the preferred models have a halo mass range lower
than the minimum of $M_{\rm min}<10^{11.4}$\hMsun considered in
abundance matching at $z=3$. Thus confirming the expectations of LAEs
to be found in less massive halos than those for LBGs.

A detailed analysis of the spectral and photometric properties of LAEs
coupled to the kind of analysis performed in this paper can be a guide
in the study of the properties of low mass dark matter halos at
$z=3.1$, extending the capabilities of abundance matching methods.

\subsection{Caveats of our method}

There are two important caveats for the work presented here. The first is the
assumption of a single LAE per dark matter halo. It is expected
for sub-halos in the simulation to host satellited galaxies. However,
it has been found in analysis based on the shape of the correlation
function [citation needed] that satellite galaxies are not a dominant
population, making our initial approximation a reasonable one.

The second caveat are the precise values for the mass intervals. These
numbers are defined from the halos defined in using the BDM halo
finder with a density threshold of 200 times the critical
density. Different halo finders and definitions for the detection
density threshold can yield different masses up to a factor of 2
[citation needed].

@@

In this \documentname we look for constraints on the preferred mass
and occupation fraction of dark matter halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We perform this study
paying attention to the impact of cosmic variance in these results. We
build a large number of mock catalogs matching observational
geometries. The mocks are constructed from a N-body simulation
following a simple recipe to assign a single LAE to each halo. Only
a fraction $f_{\rm occ}$ of halos with a mass range  $M_{\rm
  min}<M_{\rm h}<M_{\rm   max}$ can host a LAE. We proceed with a
thorough exploration of the space of free parameters $M_{\rm min},
M_{\rm max}$ and $f_{\rm occ}$ to find mocks that are consistent with
two observational constraints: the surface number density and the
angular correlation function. Out of the initial $9000$ 
combinations of parameters in the model we find $49$ arrangements
consistent with observations.

The successful parameter arrangements present a wide range of possibilities for
DM hosting LAE hosts at $z=3.1$. This means that the available LAEs' spatial
information is not sufficient to put a tight constraint on the mass
and occupation fraction of DM halos hosting LAEs. In particular it is
not possible to put a strong constraint on the occupation
fraction. Such a wide range in solutions is facilitated by the large
dispersion in the statistics derived from the mocks. All the halo mass
ranges and occupation fractions deduced in previous analysis in
previous analysis \citep[i.e.][]{Gawiser2007,Ouchi2010} are a subset of the
models we find in this \documentname.

Nevertheless, the successful models can be split into three families
depending the mass range $\Delta M=\log_{10}M_{\rm max} -
\log_{10}M_{\rm min}$ and the occupation fraction $f_{\rm occ}$. The
first family is narrow both in $\Delta M$ and $f_{\rm occ}$, a second
family is only narrow in $\Delta M$ and the third is narrow in $f_{\rm
  occ}$ and broad in $\Delta M$.

We find a large number of models with a narrow mass range
$\Delta M<1.0$ dex that are consistent in every aspect with the
spatial distribution of observed LAEs. In order to have this characteristic
it is required that halos around the mass $M_{\rm   max}$ for each
model become inefficient in hosting detectable LAEs. Two conditions
can explain this. The first is having LAEs with a decreasing \ly
escape fraction with increasing mass, the second is having larger screening
effects by neutral Hydrogen around the most massive systems
\citep{Laursen2009,ForeroRomero2011}.

A third aspect that stands out is the existence of a dominant family
of models with narrow $\Delta M<1.0$, as described before, and also a
low occupation fraction $f_{\rm occ}<0.3$. This gives support to the
common wisdom stating that only a small fraction of star forming
galaxies can be detected as strong LAEs.

However, there are extreme cases with a very narrow $\Delta M<0.3$, meaning that
there is barely a factor of $\sim 2$ between the minimum and halo
mass. If these models turn out to be confirmed, this would be a great
challenge for galaxy formation models to explain how is that LAEs can
be hosted in such a narrow range of halo mass.

In summary, we show how current observations of LAEs cannot
put a tight constraint on the physical models. We  foresee that the
new observations with new instruments (such as MUSE, Hyper SuprimeCam
and HETDEX) covering larger fields and a wider range of luminosities
will be key in imposing tighter constraints on the properties of dark
matter halos hosting LAEs.

\section*{Acknowledgments} 
J.E.F-R thanks the hospitality of Changbom Park and the Korea
Institute for Advanced Study where the first full draft of this paper
was completed. The authors also thank Peter Laursen, Paulina Lira and
Alvaro Orsi for helpful comments on the physical interpretation and
presentation of our results.
